---
title: "spatial_data_analysis"
author: "dhurv_jani"
date: "2024-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

# removes all the objects stored in the current workspace 
rm(list = ls())

```



## Install the required packages.


``` {r}
library(maps)
library(sf)
library(terra)
library(tidyterra)
library(classInt)
library(ggplot2)
library(RColorBrewer)
```


## Outlines

```{r}

# world_sf
world_sf <- st_as_sf(maps::map("world", plot = FALSE, fill = TRUE))
world_otl_sf <- st_geometry(world_sf)
plot(world_otl_sf)

```


```{r}

# conus_sf
conus_sf <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
conus_otl_sf <- st_geometry(conus_sf)
plot(conus_otl_sf)

```

## Temperature in California

```{r}
if (!require("rspat")) remotes::install_github('rspatial/rspat')
```


```{r}
library("remotes")
library("rspat")
```


## Load the data

```{r}

library(rspat)
d <- spat_data('precipitation')
head(d)

```


## Compute annual precipitation

```{r}

mnts <- toupper(month.abb)
d$prec <- rowSums(d[, mnts])
plot(sort(d$prec), ylab="Annual precipitation (mm)", las=1, xlab="Stations")

```


## Plotting the map


```{r}

dsp <- vect(d, c("LONG", "LAT"), crs="+proj=longlat +datum=NAD83")
CA <- spat_data("counties")

# define groups for mapping
cuts <- c(0,200,300,500,1000,3000)

# set up a palette of interpolated colors
blues <- colorRampPalette(c('yellow', 'orange', 'blue', 'dark blue'))
plot(CA, col="light gray", lwd=4, border="dark gray")
plot(dsp, "prec", type="interval", col=blues(10), legend=TRUE, cex=2,
    breaks=cuts, add=TRUE, plg=list(x=-117.27, y=41.54))
lines(CA)

```


## Transform longitude/latitude to planar coordinates, using the commonly used coordinate reference system for California (“Teale Albers”) to assure that our interpolation results will align with other data sets we have.

```{r}

TA <- "+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +datum=WGS84 +units=m"
dta <- project(dsp, TA)
cata <- project(CA, TA)

```




## Null Model

```{r}

RMSE <- function(observed, predicted) {
  sqrt(mean((predicted - observed)^2, na.rm=TRUE))
}

```



## Get the RMSE for the Null-model

```{r}
null <- RMSE(mean(dsp$prec), dsp$prec)
null
```
So 435 is our target. Can we do better (have a smaller RMSE)?



## Proximity Polygons

```{r}

v <- voronoi(dta)
plot(v)
points(dta)

```


## Let’s cut out what is not California, and map precipitation.

```{r}
vca <- crop(v, cata)
plot(vca, "prec")
```


## Now, we can $rasterize$ the results in the following way:

```{r}
r <- rast(vca, res=10000)
vr <- rasterize(vca, r, "prec")
plot(vr)

```

## Use 5-fold cross-validation to evaluate this model.

```{r}

set.seed(5132015)
kf <- sample(1:5, nrow(dta), replace=TRUE)
rmse <- rep(NA, 5)
for (k in 1:5) {
  test <- dta[kf == k, ]
  train <- dta[kf != k, ]
  v <- voronoi(train)
  p <- extract(v, test)
  rmse[k] <- RMSE(test$prec, p$prec)
}
rmse
mean(rmse)

# relative model performance
perf <- 1 - (mean(rmse) / null)
round(perf, 3)


```


## Nearest Neighbours Interpolation

```{r}
library("gstat")
```

```{r}

library(gstat)
d <- data.frame(geom(dta)[,c("x", "y")], as.data.frame(dta))
head(d)


gs <- gstat(formula=prec~1, locations=~x+y, data=d, nmax=5, set=list(idp = 0))
nn <- interpolate(r, gs, debug.level=0)
nnmsk <- mask(nn, vr)
plot(nnmsk, 1)

```

## Again cross-validate the results:

```{r}

rmsenn <- rep(NA, 5)
for (k in 1:5) {
  test <- d[kf == k, ]
  train <- d[kf != k, ]
  gscv <- gstat(formula=prec~1, locations=~x+y, data=train, nmax=5, set=list(idp = 0))
  p <- predict(gscv, test, debug.level=0)$var1.pred
  rmsenn[k] <- RMSE(test$prec, p)
}
rmsenn


mean(rmsenn)

1 - (mean(rmsenn) / null)
```

## Inverse Distance Weighted

```{r}

library(gstat)
gs <- gstat(formula=prec~1, locations=~x+y, data=d)
idw <- interpolate(r, gs, debug.level=0)
idwr <- mask(idw, vr)
plot(idwr, 1)

```


## Question 4: IDW generated rasters tend to have a noticeable artefact. What is that and what causes that?

## Cross-validate again. We can use predict for the locations of the test points

```{r}

rmse <- rep(NA, 5)
for (k in 1:5) {
  test <- d[kf == k, ]
  train <- d[kf != k, ]
  gs <- gstat(formula=prec~1, locations=~x+y, data=train)
  p <- predict(gs, test, debug.level=0)
  rmse[k] <- RMSE(test$prec, p$var1.pred)
}
rmse

mean(rmse)

1 - (mean(rmse) / null)

```















